Chemical Tagging Using Minimum Message Length. I.\\ Gaussian Mixture Modelling

### 2017-04-20 articles/chemical-tagging-gmm/ms.tex
enough rough stuff on the intro

\subsection{What do we want to infer from chemical abundances?}
\label{section:the-problem}

@@

There has been substantial progress on chemical tagging, despite its difficulty.

% Most have opted to address individual components of what is listed above,
% in idealistic situations.

% Progress towards Chemical Tagging

% - Mitschang for a euclidian distance in terms of seeing how different stars are

% - Dimensionality of chemical abundance space

% - Chemical homogeniety of clusters

% - Simple clustering of abundances

% - Dopplegangers

% -

Previous approaches have tried to solve separate components of chemical tagging (e.g. clustering). Some notes on this.

These approaches are not consistent, because the problems are intertwined (e.g., clustering, latent factors).

% What work has been done on chemical tagging?

% Unfortunately these approaches have only sought to solve individual problems
% within chemical tagging. Doing so -- and then later using those results to
% infer something else about the galaxy -- makes both results inconsistent,
% both statistically, and otherwise.

In Section \ref{sec:model} we introduce a simplistic approach to chemical
tagging that relies on probabilistically modelling stellar chemical abundances
as a mixture of (potentially infinite) multivariate Gaussian distributions.

In this \article\ we will introduce a simple probabilistic approach to
chemical tagging, where the data are modelled as a mixture of multivariate
Gaussian distributions (Section \ref{sec:the-model}). In Section \ref{sec:mml}
we introduce the principle of minimum message length (MML), and formulate our
objective function by minimising the total message length. Our model is
simplistic in the sense that we will ignore many of the self-consistency
issues already introduced. However, in Section \ref{sec:experiments} we show
that our MML-based approach is superior to all other methods considered for
chemical tagging, all of which cannot alleviate the self-consistency issues.
In Section \ref{sec:discussion} we will describe how MML can overcome these
aforementioned issues, in order to construct a fully-consistent probabilistic
approach to chemical tagging.

>>>>

\subsection{What do we want to infer from chemical abundances?}
\label{sec:the-problem}

@@

There has been substantial progress on chemical tagging in the last decade.
Given the difficulty of the problem, most studies have opted to address one
(or a few) of the challenges listed in Section~\ref{sec:the-problem}. 
The most common issue addressed is group-finding (items 1-4). 
Some works have sought to construct a similarity metric \citep{Mitschang}
based on chemical abundances of undisrupted clusters. 
Others have applied group-finding algorithms \citep{Hogg,Gregor}.

- Dimensionality of chemical abundance space.

- Chemical homogeneity.

- Dopplegangers.

The aforementioned chemical tagging experiments have sought to address only 
individual challenges associated with chemical tagging.
These challenges are all related, and therefore addressing them separately
will yield results that are not self-consistent, and do not make use of all
the information available.

In this \article\ we will introduce a simple probabilistic approach to
chemical tagging, where the data are modelled as a mixture of multivariate
Gaussian distributions and the number of mixtures is unknown (Section 
\ref{sec:the-model}). 
In Section \ref{sec:mml} we introduce the principle of minimum message length
(MML) and formulate our objective function.
Although our model is simplistic in the sense that we will only address some
aspects of chemical tagging, in Section \ref{sec:experiments} our experiments
show that our MML-based approach is superior to all other methods considered 
for chemical tagging.
In Section \ref{sec:discussion} we describe how the MML principle can be used
to construct a fully consistent probabilistic approach to chemical tagging,
and simultaneously address the challenges outlined in Section 
\ref{sec:the-problem}.
